---
title: "EDA"
author: "xj2249"
date: "2019/12/4"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(DataExplorer)
library(broom)
library(arsenal)
```

# clean data
```{r}
law_df <- 
    read_csv("Lawsuit.csv") %>% 
    janitor::clean_names() %>% 
    select(-id) %>% 
    mutate(gender = factor(gender,labels = c("Female","Male")),
           gender = relevel(gender, ref = "Male"),
           dept = factor(dept, labels = c("bio","phy","gene","pedia","med","sur")),
           clin = factor(clin, labels = c("research","clinical")),
           cert = factor(cert, labels = c("not certified","certified")),
           rank = factor(rank,labels = c("assistant","associate","full")),
           ave_sal = (sal94 + sal95)/2
           )
```

## consider transform response variable
```{r}
law_df %>% 
    ggplot(aes(x = ave_sal, y = ..density..)) +
    geom_histogram() + 
    labs(title = "Distribution of salary", x = "salary")

law_df %>% 
    ggplot(aes(sample = ave_sal)) + 
    stat_qq() +
    stat_qq_line() + 
    labs(title = "Normal Q-Q plot" , x = "Theoretical Quantiles", y = "Sample Quantiles")

law_df <- 
    law_df %>% 
    mutate(log_ave_sal = log(ave_sal))

law_df %>% 
    ggplot(aes(sample = log_ave_sal)) + 
    stat_qq() +
    stat_qq_line() + 
    labs(title = "Normal Q-Q plot" , x = "Theoretical Quantiles", y = "Sample Quantiles")
```


# 1. Confounders 
## 1.1 table1(covariates and outcome by gender)
```{r}
# descriptive statistics for variables of interest
control_table <- tableby.control(
        total = T,
        test = F,
        numeric.stats = c("meansd","medianq1q3","range"),
        stats.labels = list(meansd = "Mean (SD)",
                            medianq1q3 = "Median (Q1, Q3)",
                            range = "Min - Max"),
        digits = 2
        ) 
# cat.stats = "countrowpct" 

law_df %>% 
        tableby(gender ~ .,
                data = .,
                control = control_table) %>% 
        summary(text = TRUE) %>% 
        kableExtra::kable(caption = "Characcteristics of college faculty") 
```

The `clin` and `cert` has a p-value >0.05 (but < 0.20), suggesting that the composition of  `clin` and `cert` are not very different in female and male.
But still, we will check their impact on coefficient of female later.

# 1.2 exploratory visualization (relationship between salary and exposure/covariate)
## 1.2.1 categorical
```{r}
map(law_df[,c(1:4,7)], 
    ~ggplot(aes(y = log_ave_sal,x = .x, fill = gender),data = law_df) + geom_boxplot()) %>% 
    cowplot::plot_grid(plotlist = .,ncol = 2,labels = colnames(law_df[,c(1:4,7)]))
```

I want yo draw your attention to the `rank` plot. 
Seems that for male, salary doesn't change that much from assistant to full professor. But for female, it changes a lot.   

## 1.2.2 continuous 
```{r}
map(law_df[,5:6], 
    ~ggplot(aes(y = log_ave_sal,x = .x, color = gender),data = law_df) + 
            geom_point() + 
            geom_smooth(method = "lm",se = F)) %>%
    cowplot::plot_grid(plotlist = .,ncol = 1,labels = colnames(law_df[,5:6]))
```

We see a non-parallel(slightly) slope of `exper`, so we may consider interaction term in our model. 

## 1.3 Adjusted vs crude coefficients of female.
```{r}
crude <- 
    lm(log_ave_sal ~ gender, data = law_df) %>% 
    broom::tidy() %>% 
    filter(term == "genderFemale") %>% 
    pull(estimate)

adj <-
    law_df %>% 
    select(dept,clin,cert,prate,exper,rank) %>% 
    map(~lm(log_ave_sal ~ gender + .x, data = law_df) %>% broom::tidy()) %>% 
    map_dbl(~filter(.x,term == "genderFemale") %>% pull(estimate)) 

change <- round(100*(crude - adj)/crude,digits = 2)
change
```

Rule of thumb, these covariate can all be confounders. (?)

# 2. Interaction:only consider two-way for now
First, I think it's ok(or not wrong) not to include interaction in our model. 
But's it's interesting to explore the interaction

# 3.Multiple linear regression model
## 3.1 without any interaction
```{r}
full <- lm(log_ave_sal ~ gender + dept + clin + cert + prate + exper + rank, data = law_df) 
summary(full)
par(mfrow = c(2,2))
plot(full)
HH::vif(full)
```

Ah. May be... there is no "obvious" discrimination. (coz these confouders maybe actually mediators.)

## stepwise(not sure if necessary) 
```{r}
lower_gender <- lm(log_ave_sal ~ gender, data = law_df) 
step(full,scope = list(lower = lower_gender))
model_stepwise <- lm(log_ave_sal ~ gender + dept + clin + cert + exper + rank, data = law_df) 
summary(model_stepwise )
```

## 3.2  with some interaction
```{r}
inter1 <- lm(log_ave_sal ~ gender + dept +clin + cert  +  gender*exper + gender*rank,data = law_df) 
summary(inter1 )

inter2 <- lm(log_ave_sal ~ gender + dept +clin + cert  +  gender*exper + rank,data = law_df) 
summary(inter2)

```






