---
title: "model"
author: "Qi Yuchen"
date: "2019/12/9"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(faraway)
library(broom)
library(dplyr)
library(HH)
library(leaps)
library(caret)
library(purrr)
library(DataExplorer)
library(broom)
library(arsenal)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

# Data cleaning

```{r}
# import
df = read_csv("Lawsuit.csv") %>% 
  janitor::clean_names() %>% 
  mutate(dept = factor(dept, levels = c(1:6), labels = c("Biochemistry/Molecular Biology", "Physiology ", "Genetics", "Pediatrics", "Medicine", "Surgery")),
         gender = factor(gender, levels = c(0, 1), labels = c("female","male")),
         clin = factor(clin, levels = c(1, 0), labels = c("clinical","research")),
         cert = factor(cert, levels = c(1, 0), labels = c("certified","not_certified")),
         rank = factor(rank, levels = c(1, 2, 3), labels = c("assistant", "associate", "full")),
         sal = (sal94 + sal95)/2)

# consider sal94 only first 
df_94 = df %>%   
  mutate(log_sal94 = log(sal94)) %>% 
  dplyr::select(-sal95, -id, -sal94, -sal)

# from EDA
law_df <- 
    read_csv("Lawsuit.csv") %>% 
    janitor::clean_names() %>% 
    dplyr::select(-id) %>% 
    mutate(gender = factor(gender,labels = c("Female","Male")),
           gender = relevel(gender, ref = "Male"),
           dept = factor(dept, labels = c("bio","phy","gene","pedia","med","sur")),
           clin = factor(clin, labels = c("research","clinical")),
           cert = factor(cert, labels = c("not certified","certified")),
           rank = factor(rank,labels = c("assistant","associate","full")),
           )
```

# Model

## main effects

### Confounders

First consider confounders

Adjusted vs crude coefficients of female

```{r}
# form EDA
crude <- 
    lm(sal94 ~ gender, data = law_df) %>% 
    broom::tidy() %>% 
    filter(term == "genderFemale") %>% 
    pull(estimate)

adj <-
    law_df %>% 
    dplyr::select(dept,clin,cert,prate,exper,rank) %>% 
    map(~lm(sal94 ~ gender + .x, data = law_df) %>% broom::tidy()) %>% 
    map_dbl(~filter(.x,term == "genderFemale") %>% pull(estimate)) 

change <- round(100*(crude - adj)/crude,digits = 2)
change

```

Rule of thumb, these covariate can all be confounders.

Then inlcude them all in the original model.

### model selection process

stepwise

```{r}
fit_94 <- lm(log_sal94 ~ gender, data = df_94)

biggest <- formula(lm(log_sal94 ~ ., df_94))

step(fit_94, direction = 'both', scope = biggest)
```

Stepwise using AIC does not include gender.

Criteria

```{r}
best_1 <- function(model, ...) 
{
  subsets <- regsubsets(formula(model), model.frame(model), ...)
  subsets <- with(summary(subsets),
                  cbind(p = as.numeric(rownames(which)), which, rss, rsq, adjr2, cp, bic))
  
  return(subsets)
}  


full = lm(log_sal94 ~ ., df_94)

best_1(full) 

```

So is using criteria.

### Use VIF to see collinearity

```{r}
vif(full)
```

prate's VIF is more than 10, we may drop it.

```{r}
fit = lm(log_sal94 ~gender+dept+clin+cert+exper+rank, data = df_94)

vif(fit)
```

Now each VIF < 5, which is good.

# interaction term

```{r}
## general test

# no prate
fit_conf = lm(log_sal94 ~gender+dept+clin+cert+exper+rank, data = df_94)

fit_int = lm(log_sal94 ~gender+dept+clin+cert+exper+gender*dept+gender*clin+gender*cert+gender*exper+gender*rank, data = df_94)

anova(fit_conf, fit_int)
```

Maybe no interaction term.

The model now is log_sal94 ~gender+dept+clin+cert+exper+rank

```{r}
summary(fit)
```

Adjusted R squared is 0.93, indicating that the model fits well.

From this model, there is no "obvious" discrimination. However, there may be mediators.

# stratified regression

```{r}
stratified_dept = df_94 %>%
  group_by(dept) %>%
  summarize(
      n = n(),
      coef =  lm(log_sal94 ~ gender)$coef["gendermale"]
            )
stratified_dept %>% 
    knitr::kable()

stratified_clin = df_94 %>%
  group_by(clin) %>%
  summarize(
      n = n(),
      coef =  lm(log_sal94 ~ gender)$coef["gendermale"]
            )
stratified_clin %>% 
    knitr::kable()

stratified_cert = df_94 %>%
  group_by(cert) %>%
  summarize(
      n = n(),
      coef =  lm(log_sal94 ~ gender)$coef["gendermale"]
            )
stratified_cert %>% 
    knitr::kable()

stratified_rank = df_94 %>%
  group_by(rank) %>%
  summarize(
      n = n(),
      coef =  lm(log_sal94 ~ gender)$coef["gendermale"]
            )
stratified_rank %>% 
    knitr::kable()
```

