---
title: "results"
author: "Qi Yuchen"
date: "2019/12/11"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(faraway)
library(broom)
library(dplyr)
library(HH)
library(leaps)
library(caret)
library(purrr)
library(DataExplorer)
library(broom)
library(arsenal)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

# Data cleaning

```{r}
# import
df = read_csv("Lawsuit.csv") %>% 
  janitor::clean_names() %>% 
  mutate(dept = factor(dept, levels = c(1:6), labels = c("Biochemistry/Molecular Biology", "Physiology ", "Genetics", "Pediatrics", "Medicine", "Surgery")),
         gender = factor(gender, levels = c(0, 1), labels = c("female","male")),
         clin = factor(clin, levels = c(1, 0), labels = c("clinical","research")),
         cert = factor(cert, levels = c(1, 0), labels = c("certified","not_certified")),
         rank = factor(rank, levels = c(1, 2, 3), labels = c("assistant", "associate", "full")),
         sal = (sal94 + sal95)/2)

# consider log mean sal only 
df_sal = df %>%   
  mutate(log_sal = log(sal)) %>% 
  dplyr::select(-sal95, -id, -sal94, -sal)
```


# Data exploration

## distribution of the outcome


From the distribution of average salary we applied log transformation to the response variable.

## table1(covariates and outcome by gender)


## visualizing the relationship between response variable and predictor variables by gender

see EDA





# Model

## main effects

### Confounders

First we consider confounders.

We check the adjusted and crude coefficients of gendermale.

```{r}
crude <- 
    lm(log_sal ~ gender, data = df_sal) %>% 
    broom::tidy() %>% 
    filter(term == "gendermale") %>% 
    pull(estimate)

adj <-
    df_sal %>% 
    dplyr::select(dept,clin,cert,prate,exper,rank) %>% 
    map(~lm(log_sal ~ gender + .x, data = df_sal) %>% broom::tidy()) %>% 
    map_dbl(~filter(.x,term == "gendermale") %>% pull(estimate)) 

change <- round(100*(crude - adj)/crude,digits = 2)
change

```

As the magtitude of each coefficient change is around or larger than 10%, all covariates may be confounders.Then we inlcude them all in the model.

### Use VIF to see collinearity

```{r}


full = lm(log_sal ~ ., data = df_sal)

vif(full)
```

Since prate's VIF is more than 10, we may drop it.

```{r}
fit = lm(log_sal ~gender+dept+clin+cert+exper+rank, data = df_sal)

vif(fit)

anova(fit, full)
```

Now each VIF < 5, which is good. F test also indicates no prate.

### Criteria

```{r}
best_1 <- function(model, ...) 
{
  subsets <- regsubsets(formula(model), model.frame(model), method = "exhaustive", nvmax = NULL, force.in = "gendermale",  ...)
  subsets <- with(summary(subsets),
                  cbind(p = as.numeric(rownames(which)), which, rss, rsq, adjr2, cp, bic))
  
  return(subsets)
}  


best_1(full) 

subsets = regsubsets(formula(full), model.frame(full), method = "exhaustive", nvmax = NULL, force.in = "gendermale")

# add interaction

fit_int = lm(log_sal ~gender+dept+clin+cert+exper+prate+gender*dept+gender*clin+gender*cert+gender*exper+gender*rank+gender*prate, data = df_sal)
best_1(fit_int)

fit_optimal = lm(log_sal ~gender+dept+clin+cert+exper+gender*clin+gender*exper, data = df_sal)
summary(fit_optimal)
```

So is using criteria. If force in gender, the best is without prate. clin and exper interaction will be selected if add interaction terms.


## interaction term (not sure whether to test)

```{r}
## general test 

# no prate
fit_conf = lm(log_sal ~gender+dept+clin+cert+exper+rank, data = df_sal)
summary(fit_conf)

fit_int = lm(log_sal ~gender+dept+clin+cert+exper+rank+gender*exper, data = df_sal)
summary(fit_int)

anova(fit_conf, fit_int)

```

Interaction term $gender*exper$ is significant, thus we may conside it in our model.


## stratified regression

```{r}
stratified_dept = df_sal %>%
  group_by(dept) %>%
  summarize(
      n = n(),
      coef =  lm(log_sal ~ gender+clin+cert+exper+rank)$coef["gendermale"],
      p = summary(lm(log_sal ~ gender+clin+cert+exper+rank))$coefficients["gendermale",4]
            )
stratified_dept %>% 
    knitr::kable()

stratified_clin = df_sal %>%
  group_by(clin) %>%
  summarize(
      n = n(),
      coef =  lm(log_sal ~ gender + dept + cert + exper + 
    rank)$coef["gendermale"],
      p = summary(lm(log_sal ~ gender + dept + cert + exper + 
    rank))$coefficients["gendermale",4]
            )
stratified_clin %>% 
    knitr::kable()

stratified_cert = df_sal %>%
  group_by(cert) %>%
  summarize(
      n = n(),
      coef =  lm(log_sal ~ gender + dept + clin + exper + 
    rank)$coef["gendermale"],
      p = summary(lm(log_sal ~ gender + dept + clin + exper + 
    rank))$coefficients["gendermale",4]
            )
stratified_cert %>% 
    knitr::kable()

stratified_rank = df_sal %>%
  group_by(rank) %>%
  summarize(
      n = n(),
      coef =  lm(log_sal ~ gender + dept + clin + cert + exper)$coef["gendermale"],
      p = summary(lm(log_sal ~ gender + dept + clin + cert + exper))$coefficients["gendermale",4]
            )
stratified_rank %>% 
    knitr::kable()
```


```{r}
df_exper = df_sal %>%
  mutate(exper_fct = case_when(
      exper < 6 ~ "0",
      exper >= 6 & exper < 9 ~ "1",
      exper >= 9 & exper < 14 ~ "2",
      exper >= 14 ~ "3",
      TRUE ~ ""
  )) %>% 
    mutate(exper = factor(exper_fct)) %>% 
    dplyr::select(-exper_fct)

stratified_exper = df_exper %>%
  group_by(exper) %>%
  summarize(
      n = n(),
      coef =  lm(log_sal ~ gender + dept + clin + cert + rank)$coef["gendermale"],
      p = summary(lm(log_sal ~ gender + dept + clin + cert + rank))$coefficients["gendermale",4]
            )
stratified_exper %>% 
    knitr::kable()

df_exper  %>%
  group_by(exper, rank) %>%
  summarize(
      n = n())
```



# Model diagnostics

```{r}
plot(fit_int)
plot(lm(log_sal ~gender+dept+clin+cert+exper+rank*gender+gender*exper, data = df_sal))

stu_res<-rstandard(fit_int)
stu_res[abs(stu_res)>2.5]

influence.measures(fit_int) %>% 
  summary()


df_sal_noinflu = df_sal[-184, ]
df_sal_noinflu = df_sal_noinflu[-58, ]
df_sal_noinflu = df_sal_noinflu[-216, ]
plot(lm(log_sal ~gender+dept+clin+cert+exper+rank*gender+gender*exper, data = df_sal_noinflu))
temp = lm(log_sal ~gender+dept+clin+cert+exper+rank*gender+gender*exper, data = df_sal_noinflu)
summary(temp)

df_exper_noinflu = df_sal_noinflu %>%
  mutate(exper_fct = case_when(
      exper < 6 ~ "0",
      exper >= 6 & exper < 9 ~ "1",
      exper >= 9 & exper < 14 ~ "2",
      exper >= 14 ~ "3",
      TRUE ~ ""
  )) %>% 
    mutate(exper = factor(exper_fct)) %>% 
    dplyr::select(-exper_fct)

stratified_exper_noinflu = df_exper_noinflu %>%
  group_by(exper) %>%
  summarize(
      n = n(),
      coef =  lm(log_sal ~ gender + dept + clin + cert + rank)$coef["gendermale"],
      p = summary(lm(log_sal ~ gender + dept + clin + cert + rank))$coefficients["gendermale",4]
            )
stratified_exper %>% 
    knitr::kable()
```

