---
title: "results"
author: "Qi Yuchen"
date: "2019/12/11"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(faraway)
library(broom)
library(dplyr)
library(HH)
library(leaps)
library(caret)
library(purrr)
library(DataExplorer)
library(broom)
library(arsenal)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

# Data cleaning

```{r}
# import
df = read_csv("Lawsuit.csv") %>% 
  janitor::clean_names() %>% 
  mutate(dept = factor(dept, levels = c(1:6), labels = c("Biochemistry/Molecular Biology", "Physiology ", "Genetics", "Pediatrics", "Medicine", "Surgery")),
         gender = factor(gender, levels = c(0, 1), labels = c("female","male")),
         clin = factor(clin, levels = c(1, 0), labels = c("clinical","research")),
         cert = factor(cert, levels = c(1, 0), labels = c("certified","not_certified")),
         rank = factor(rank, levels = c(1, 2, 3), labels = c("assistant", "associate", "full")),
         sal = (sal94 + sal95)/2)

# consider log mean sal only 
df_sal = df %>%   
  mutate(log_sal = log(sal)) %>% 
  dplyr::select(-sal95, -id, -sal94, -sal)
```


# Data exploration

## distribution of the outcome

From the distribution of average salary we applied log transformation to the response variable.

## table1(covariates and outcome by gender)

## visualizing the relationship between response variable and predictor variables by gender

see EDA





# Model

## main effects

### Confounders

First we consider confounders.

We check the adjusted and crude coefficients of gendermale.

```{r}
crude <- 
    lm(log_sal ~ gender, data = df_sal) %>% 
    broom::tidy() %>% 
    filter(term == "gendermale") %>% 
    pull(estimate)

adj <-
    df_sal %>% 
    dplyr::select(dept,clin,cert,prate,exper,rank) %>% 
    map(~lm(log_sal ~ gender + .x, data = df_sal) %>% broom::tidy()) %>% 
    map_dbl(~filter(.x,term == "gendermale") %>% pull(estimate)) 

change <- round(100*(crude - adj)/crude,digits = 2)
change

```

As the magtitude of each coefficient change is around or larger than 10%, all covariates may be confounders.Then we inlcude them all in the model.

### Use VIF to see collinearity

```{r}
full = lm(log_sal ~ ., data = df_sal)

vif(full)
```

Since prate's VIF is more than 10, we may drop it.

```{r}
fit = lm(log_sal ~gender+dept+clin+cert+exper+rank, data = df_sal)

vif(fit)

anova(fit, full)
```

Now each VIF < 5, which is good. F test also indicates no prate.

### Criteria

```{r}
best_1 <- function(model, ...) 
{
  subsets <- regsubsets(formula(model), model.frame(model), method = "exhaustive", nvmax = NULL, force.in = "gendermale",  ...)
  subsets <- with(summary(subsets),
                  cbind(p = as.numeric(rownames(which)), which, rss, rsq, adjr2, cp, bic))
  
  return(subsets)
}  


best_1(full) 

subsets = regsubsets(formula(full), model.frame(full), method = "exhaustive", nvmax = NULL, force.in = "gendermale")

# add interaction

fit_int = lm(log_sal ~gender+dept+clin+cert+exper+prate+gender*dept+gender*clin+gender*cert+gender*exper+gender*rank+gender*prate, data = df_sal)
best_1(fit_int)

fit_optimal = lm(log_sal ~gender+dept+clin+cert+exper+gender*clin+gender*exper, data = df_sal)
summary(fit_optimal)
```

So is using criteria. If force in gender, the best is without prate. clin and exper interaction will be selected if add interaction terms.


## interaction term (not sure whether to test)

```{r}
## general test 

# no prate
fit_conf = lm(log_sal ~gender+dept+clin+cert+exper+rank, data = df_sal)
summary(fit_conf)

fit_int = lm(log_sal ~gender+dept+clin+cert+exper+rank+gender*exper, data = df_sal)
summary(fit_int)

anova(fit_conf, fit_int)

```

Interaction term $gender*exper$ is significant, thus we may conside it in our model.


## stratified regression

```{r}
stratified_dept = df_sal %>%
  group_by(dept) %>%
  summarize(
      n = n(),
      coef =  lm(log_sal ~ gender+clin+cert+exper+rank)$coef["gendermale"],
      p = summary(lm(log_sal ~ gender+clin+cert+exper+rank))$coefficients["gendermale",4]
            )
stratified_dept %>% 
    knitr::kable()

stratified_clin = df_sal %>%
  group_by(clin) %>%
  summarize(
      n = n(),
      coef =  lm(log_sal ~ gender + dept + cert + exper + 
    rank)$coef["gendermale"],
      p = summary(lm(log_sal ~ gender + dept + cert + exper + 
    rank))$coefficients["gendermale",4]
            )
stratified_clin %>% 
    knitr::kable()

stratified_cert = df_sal %>%
  group_by(cert) %>%
  summarize(
      n = n(),
      coef =  lm(log_sal ~ gender + dept + clin + exper + 
    rank)$coef["gendermale"],
      p = summary(lm(log_sal ~ gender + dept + clin + exper + 
    rank))$coefficients["gendermale",4]
            )
stratified_cert %>% 
    knitr::kable()

stratified_rank = df_sal %>%
  group_by(rank) %>%
  summarize(
      n = n(),
      coef =  lm(log_sal ~ gender + dept + clin + cert + exper)$coef["gendermale"],
      p = summary(lm(log_sal ~ gender + dept + clin + cert + exper))$coefficients["gendermale",4]
            )
stratified_rank %>% 
    knitr::kable()
```


```{r}
df_exper = df_sal %>%
  mutate(exper_fct = case_when(
      exper < 6 ~ "0",
      exper >= 6 & exper < 9 ~ "1",
      exper >= 9 & exper < 14 ~ "2",
      exper >= 14 ~ "3",
      TRUE ~ ""
  )) %>% 
    mutate(exper = factor(exper_fct)) %>% 
    dplyr::select(-exper_fct)

stratified_exper = df_exper %>%
  group_by(exper) %>%
  summarize(
      n = n(),
      coef =  lm(log_sal ~ gender + dept + clin + cert + rank)$coef["gendermale"],
      p = summary(lm(log_sal ~ gender + dept + clin + cert + rank))$coefficients["gendermale",4]
            )
stratified_exper %>% 
    knitr::kable()

df_exper  %>%
  group_by(exper, rank) %>%
  summarize(
      n = n())
```



# Model diagnostics

```{r}
final_model = lm(log_sal ~gender+dept+clin+cert+exper+rank, data = df_sal)
par(mfrow = c(2,2))
plot(final_model)
```

# Outliers/influential points

```{r}
stu_res<-rstandard(final_model)
stu_res[abs(stu_res)>2.5]

influence.measures(final_model) %>% 
  summary()

df[184,]
```

Using studentized residuals, id 184 is an outlier in Y. using leverage values, 19 and 216 are outliers in X. Using DFFIT, 8, 184 and 216 are influential points. Using main effects only, 184 is influential.

```{r}
# consider the data without influential points
df_sal_noinflu = df_sal[-184, ]

par(mfrow = c(2,2))
plot(final_model)
temp = lm(log_sal ~gender+dept+clin+cert+exper+rank*gender+gender*exper, data = df_sal_noinflu)
summary(temp)

df_exper_noinflu = df_sal_noinflu %>%
  mutate(exper_fct = case_when(
      exper < 6 ~ "0",
      exper >= 6 & exper < 9 ~ "1",
      exper >= 9 & exper < 14 ~ "2",
      exper >= 14 ~ "3",
      TRUE ~ ""
  )) %>% 
    mutate(exper = factor(exper_fct)) %>% 
    dplyr::select(-exper_fct)

stratified_exper_noinflu = df_exper_noinflu %>%
  group_by(exper) %>%
  summarize(
      n = n(),
      coef =  lm(log_sal ~ gender + dept + clin + cert + rank)$coef["gendermale"],
      p = summary(lm(log_sal ~ gender + dept + clin + cert + rank))$coefficients["gendermale",4]
            )
stratified_exper_noinflu %>% 
    knitr::kable()

stratified_rank_noinflu = df_sal_noinflu %>%
  group_by(rank) %>%
  summarize(
      n = n(),
      coef =  lm(log_sal ~ gender + dept + clin + cert + exper)$coef["gendermale"],
      p = summary(lm(log_sal ~ gender + dept + clin + cert + exper))$coefficients["gendermale",4]
            )
stratified_rank_noinflu %>% 
    knitr::kable()
```

Not significant now, -184 usinf main effects model or -216, -184, -8 using interaction model.
