---
title: "results"
author: "Qi Yuchen"
date: "2019/12/11"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(faraway)
library(broom)
library(dplyr)
library(HH)
library(leaps)
library(caret)
library(purrr)
library(DataExplorer)
library(broom)
library(arsenal)
library(ggpubr)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom") + theme(plot.title = element_text(hjust = 0.5)))

```

# Data cleaning

```{r}
# import
df = read_csv("Lawsuit.csv") %>% 
  janitor::clean_names() %>% 
  mutate(dept = factor(dept, levels = c(1:6), labels = c("Biochemistry/Molecular Biology", "Physiology ", "Genetics", "Pediatrics", "Medicine", "Surgery")),
         gender = factor(gender, levels = c(0, 1), labels = c("Female","Male")),
         clin = factor(clin, levels = c(1, 0), labels = c("Clinical","Research")),
         cert = factor(cert, levels = c(1, 0), labels = c("Certified","Not certified")),
         rank = factor(rank, levels = c(1, 2, 3), labels = c("Assistant", "Associate", "Full professor")),
         sal = (sal94 + sal95)/2) 

# consider log mean sal only 
df_sal = df %>%   
  mutate(log_sal = log(sal)) %>% 
  dplyr::select(-sal95, -id, -sal94, -sal)
```


# Data exploration

## distribution of the outcome
```{r}
hist_ave <-
    df %>% 
    ggplot(aes(x = sal, y = ..density..)) +
    geom_histogram() + 
    labs(title = "Distribution of Average Salary", x = "Average Salary" , y = "Density")

qq_ave <-
    df %>% 
    ggplot(aes(sample = sal)) + 
    stat_qq() +
    stat_qq_line() + 
    labs(title = "Normal Q-Q plot of Average Salary" , x = "Theoretical Quantiles", y = "Sample Quantiles")

hist_trans <-
    df_sal %>% 
    ggplot(aes(x = log_sal, y = ..density..)) +
    geom_histogram() + 
    labs(title = "Distribution of Transformed Salary", x = "Transformed Salary", y = "Density")


qq_trans <-
    df_sal  %>% 
    ggplot(aes(sample = log_sal)) + 
    stat_qq() +
    stat_qq_line() + 
    labs(title = "Normal Q-Q plot of Transformed Salary" , x = "Theoretical Quantiles", y = "Sample Quantiles")

ggarrange(hist_ave,hist_trans,qq_ave,qq_trans) %>% 
    annotate_figure(top = text_grob("Salary Transformation",face = "bold", size = 14))
```

From the distribution of average salary we applied log transformation to the response variable.

## table1(covariates and outcome by gender)
```{r}
# descriptive statistics for variables of interest
control_table <- tableby.control(
        total = T,
        test = F,
        numeric.stats = c("meansd","medianq1q3","range"),
        stats.labels = list(meansd = "Mean (SD)",
                            medianq1q3 = "Median (Q1, Q3)",
                            range = "Range"),
        digits = 2
        ) 

my_labels <- list(
  dept = "Department",
  clin = "Primary emphasis",
  cert = "Board certification",
  prate = "Publication rate",
  exper = "Experience",
  rank = "Rank",
  log_sal = "Log Salary"
)

tab1 <- 
    df_sal %>% 
    tableby(gender ~ .,
            data = .,
            control = control_table) %>% 
    summary(text = TRUE,labelTranslations = my_labels, title = "Summary Statistic of Data by Gender")

write2pdf(tab1, "table1.pdf")
```


## visualizing the relationship between response variable and predictor variables by gender
categorical
```{r}
dept_plot <- 
    df_sal %>% 
    ggplot(aes(y = log_sal,x = dept, fill = gender)) + 
    geom_boxplot() + 
    labs(x = "Department",y = "")

clin_plot <- 
    df_sal %>% 
    ggplot(aes(y = log_sal,x = clin, fill = gender)) + 
    geom_boxplot() + 
    labs(x = "Primary emphasis",y = "")

cert_plot <- 
    df_sal %>% 
    ggplot(aes(y = log_sal,x = cert, fill = gender)) + 
    geom_boxplot() + 
    labs(x = "Board certification", y = "")

rank_plot <- 
    df_sal %>% 
    ggplot(aes(y = log_sal,x = rank, fill = gender)) + 
    geom_boxplot() + 
    labs(x = "Rank",y = "")

gender_plot <- 
    df_sal %>% 
    ggplot(aes(y = log_sal,x = gender, fill = gender)) + 
    geom_boxplot() + 
    labs(x = "Gender",y = "")
```

continuous 
```{r}
prate_plot <-
    df_sal %>% 
    ggplot(aes(y = log_sal,x = prate, color = gender)) + 
    geom_point() + 
    geom_smooth(method = "lm",se = F) + 
    labs(title = "Salary vs Publication rate", x = "Publication rate" , y = "")


exper_plot <-
    df_sal %>% 
    ggplot(aes(y = log_sal,x = exper, color = gender)) + 
    geom_point() + 
    geom_smooth(method = "lm",se = F) +
    labs(title = "Salary vs Experience", x = "Years since obtaining MD", y = "")

figure_all <- ggarrange(dept_plot,clin_plot,cert_plot,rank_plot,gender_plot,prate_plot,exper_plot,
                    common.legend = T,legend = "bottom")

# Annotate the figure by adding a common labels
annotate_figure(figure_all,
                top = text_grob("Salary Comparison",face = "bold", size = 14),
                left = text_grob("log(salary)",rot = 90))
```

We see a non-parallel(slightly) slope of `exper`, so we may consider interaction term in our model. 

# Model

## main effects

### Confounders

First we consider confounders.

We check the adjusted and crude coefficients of gendermale.

```{r}
crude <- 
    lm(log_sal ~ gender, data = df_sal) %>% 
    broom::tidy() %>% 
    filter(term == "genderMale") %>% 
    pull(estimate)

adj <-
    df_sal %>% 
    dplyr::select(dept,clin,cert,prate,exper,rank) %>% 
    map(~lm(log_sal ~ gender + .x, data = df_sal) %>% broom::tidy()) %>% 
    map_dbl(~filter(.x,term == "genderMale") %>% pull(estimate)) 

change <- round(100*(crude - adj)/crude,digits = 2)
change 
```

As the magtitude of each coefficient change is around or larger than 10%, all covariates may be confounders.Then we inlcude them all in the model.

### Use VIF to see collinearity

```{r}
full = lm(log_sal ~ ., data = df_sal)

vif(full)
```

Since prate's VIF is more than 10, we may drop it.

```{r}
fit = lm(log_sal ~gender+dept+clin+cert+exper+rank, data = df_sal)

vif(fit)

anova(fit, full)
```

Now each VIF < 5, which is good. F test also indicates no prate.

### Criteria

```{r}
best_1 <- function(model, ...) 
{
  subsets <- regsubsets(formula(model), model.frame(model), method = "exhaustive", nvmax = NULL, force.in = "genderMale",  ...)
  subsets <- with(summary(subsets),
                  cbind(p = as.numeric(rownames(which)), which, rss, rsq, adjr2, cp, bic))
  
  return(subsets)
}  


best_1(full) 

subsets = regsubsets(formula(full), model.frame(full), method = "exhaustive", nvmax = NULL, force.in = "genderMale")

# add interaction

fit_int = lm(log_sal ~gender+dept+clin+cert+exper+prate+gender*dept+gender*clin+gender*cert+gender*exper+gender*rank+gender*prate, data = df_sal)
best_1(fit_int)

fit_optimal = lm(log_sal ~gender+dept+clin+cert+exper+gender*clin+gender*exper, data = df_sal)
summary(fit_optimal)
```

So is using criteria. If force in gender, the best is without prate. clin and exper interaction will be selected if add interaction terms.


## interaction term (not sure whether to test)

```{r results='asis'}
## general test 

# no prate
fit_conf = lm(log_sal ~gender+dept+clin+cert+exper+rank, data = df_sal)
summary(fit_conf)

fit_int = lm(log_sal ~gender+dept+clin+cert+exper+rank+gender*exper, data = df_sal)

fit_int %>% broom::tidy() %>% kableExtra::kable()

anova(fit_conf, fit_int)

```

Interaction term $gender*exper$ is significant, thus we may conside it in our model.


## stratified regression

```{r}
stratified_dept = df_sal %>%
  group_by(dept) %>%
  summarize(
      n = n(),
      coef =  lm(log_sal ~ gender+clin+cert+exper+rank)$coef["genderMale"],
      p = summary(lm(log_sal ~ gender+clin+cert+exper+rank))$coefficients["genderMale",4]
            )
stratified_dept %>% 
    knitr::kable()

stratified_clin = df_sal %>%
  group_by(clin) %>%
  summarize(
      n = n(),
      coef =  lm(log_sal ~ gender + dept + cert + exper + 
    rank)$coef["genderMale"],
      p = summary(lm(log_sal ~ gender + dept + cert + exper + 
    rank))$coefficients["genderMale",4]
            )
stratified_clin %>% 
    knitr::kable()

stratified_cert = df_sal %>%
  group_by(cert) %>%
  summarize(
      n = n(),
      coef =  lm(log_sal ~ gender + dept + clin + exper + 
    rank)$coef["genderMale"],
      p = summary(lm(log_sal ~ gender + dept + clin + exper + 
    rank))$coefficients["genderMale",4]
            )
stratified_cert %>% 
    knitr::kable()

stratified_rank = df_sal %>%
  group_by(rank) %>%
  summarize(
      n = n(),
      coef =  lm(log_sal ~ gender + dept + clin + cert + exper)$coef["genderMale"],
      p = summary(lm(log_sal ~ gender + dept + clin + cert + exper))$coefficients["genderMale",4]
            )
stratified_rank %>% 
    knitr::kable()
```


```{r}
df_exper = df_sal %>%
  mutate(exper_fct = case_when(
      exper < 6 ~ "0",
      exper >= 6 & exper < 9 ~ "1",
      exper >= 9 & exper < 14 ~ "2",
      exper >= 14 ~ "3",
      TRUE ~ ""
  )) %>% 
    mutate(exper = factor(exper_fct)) %>% 
    dplyr::select(-exper_fct)

stratified_exper = df_exper %>%
  group_by(exper) %>%
  summarize(
      n = n(),
      coef =  lm(log_sal ~ gender + dept + clin + cert + rank)$coef["genderMale"],
      p = summary(lm(log_sal ~ gender + dept + clin + cert + rank))$coefficients["genderMale",4]
            )
stratified_exper %>% 
    knitr::kable()

df_exper  %>%
  group_by(exper, rank) %>%
  summarize(
      n = n())
```



# Model diagnostics

```{r}
final_model = lm(log_sal ~gender+dept+clin+cert+exper+rank, data = df_sal)
par(mfrow = c(2,2))
plot(final_model)
```

# Outliers/influential points

```{r}
stu_res<-rstandard(final_model)
stu_res[abs(stu_res)>2.5]

influence.measures(final_model) %>% 
  summary()

df[184,]
```

Using studentized residuals, id 184 is an outlier in Y. using leverage values, 19 and 216 are outliers in X. Using DFFIT, 8, 184 and 216 are influential points. Using main effects only, 184 is influential.

```{r}
# consider the data without influential points
df_sal_noinflu = df_sal[-184, ]

par(mfrow = c(2,2))
plot(final_model)
temp = lm(log_sal ~gender+dept+clin+cert+exper+rank*gender+gender*exper, data = df_sal_noinflu)
summary(temp)

df_exper_noinflu = df_sal_noinflu %>%
  mutate(exper_fct = case_when(
      exper < 6 ~ "0",
      exper >= 6 & exper < 9 ~ "1",
      exper >= 9 & exper < 14 ~ "2",
      exper >= 14 ~ "3",
      TRUE ~ ""
  )) %>% 
    mutate(exper = factor(exper_fct)) %>% 
    dplyr::select(-exper_fct)

stratified_exper_noinflu = df_exper_noinflu %>%
  group_by(exper) %>%
  summarize(
      n = n(),
      coef =  lm(log_sal ~ gender + dept + clin + cert + rank)$coef["genderMale"],
      p = summary(lm(log_sal ~ gender + dept + clin + cert + rank))$coefficients["genderMale",4]
            )
stratified_exper_noinflu %>% 
    knitr::kable()

stratified_rank_noinflu = df_sal_noinflu %>%
  group_by(rank) %>%
  summarize(
      n = n(),
      coef =  lm(log_sal ~ gender + dept + clin + cert + exper)$coef["genderMale"],
      p = summary(lm(log_sal ~ gender + dept + clin + cert + exper))$coefficients["genderMale",4]
            )
stratified_rank_noinflu %>% 
    knitr::kable()
```

Not significant now, -184 usinf main effects model or -216, -184, -8 using interaction model.
